\phantomsection
\begin{thebibliography}{99}


\bibitem{RL}Sutton, Richard S and Barto, Andrew G \textit{Reinforcement learning: An introduction}. MIT press, 2018.

\bibitem{jordan2015}Jordan, M.I. and Mitchell, T.M. \textit{Machine learning: Trends, perspectives, and prospects}. Science, 349(6245):255--260, 2015.

\bibitem{Markov}Puterman, Martin L \textit{Markov Decision Processes: Discrete Stochastic Dynamic Programming}. John Wiley \& Sons, Inc., 1994.

\bibitem{RL2}Russell, Stuart J and Norvig, Peter \textit{Artificial intelligence: a modern approach}. MIT press, 2016.


\bibitem{thrun2005}Thrun, Sebastian and Burgard, Wolfram and Fox, Dieter \textit{Probabilistic Robotics}. MIT Press, 2005. \url{https://docs.ufpr.br/~danielsantos/ProbabilisticRobotics.pdf}


\bibitem{komorowski2018}Komorowski, Matthieu and Celi, Leo Anthony and Badawi, Omar and Gordon, Anthony C. and Faisal, A. Aldo \textit{The Artificial Intelligence Clinician learns optimal treatment strategies for sepsis in intensive care}. Nature Medicine, 24:1716–1720, 2018. \url{https://finale.seas.harvard.edu/publications/guidelines-reinforcement-learning-healthcare}



\bibitem{buehler2019}Buehler, Stefan and Ranganathan, Arun and Gupta, Rahul \textit{Reinforcement Learning for Finance}. Stanford University, 2019. \url{https://stanford.edu/~ashlearn/RLForFinanceBook/book.pdf}


\bibitem{mohri2018}Mohri, Mehryar and Rostamizadeh, Afshin and Talwalkar, Ameet \textit{Foundations of Machine Learning}. The MIT Press, 2018. \url{https://cs.nyu.edu/~mohri/mlbook/}


\bibitem{bellman1958}Bellman, Richard \textit{Dynamic Programming and Stochastic Control Processes}. Information and Control, 1:228--239, 1958.

\bibitem{sutton1988b}Sutton, Richard S. \textit{Learning to Predict by the Methods of Temporal Difference}. Machine Learning, 3:9--44, 1988.

\bibitem{watkins1989}Watkins, Christopher JCH. \textit{Learning from Delayed Rewards}. PhD Thesis, University of Cambridge, England, 1989.





\bibitem{watkins1992}Watkins, Christopher JCH and Dayan, Peter \textit{Q-learning}. Machine Learning, 8(3):279--292, 1992.




\bibitem{bdr2022}Marc G. Bellemare and Will Dabney and Mark Rowland \textit{Distributional Reinforcement Learning}. MIT Press, 2023.


\bibitem{Goodfellow-et-al-2016}Ian Goodfellow and Yoshua Bengio and Aaron Courville \textit{Deep Learning}. MIT Press, 2016. \url{http://www.deeplearningbook.org}

\bibitem{RL3}Dimitri Bertsekas \textit{Reinforcement Learning and Optimal Control}. MIT Press, 2019.



\bibitem{kober2013}Kober, Jens and Bagnell, J. Andrew and Peters, Jan \textit{Reinforcement Learning in Robotics: A Survey}. International Journal of Robotics Research, 2013. \url{https://www.ri.cmu.edu/pub_files/2013/7/Kober_IJRR_2013.pdf}

\bibitem{van2016double}Van Hasselt, Hado and Guez, Arthur and Silver, David \textit{Deep Reinforcement Learning with Double Q-learning}. Proceedings of the Thirtieth AAAI Conference on Artificial Intelligence, 2016. \url{https://www.davidsilver.uk/wp-content/uploads/2020/03/doubledqn-1.pdf}




\end{thebibliography}